---
title: "p8105_hw6_qz2493.Rmd"
author: "Qingyue Zhuo qz2493"
date: "2022-11-27"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
```

## Problem 2

### Load and Clean the Dataset

* Load the data
```{r}
raw_data = read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

* Clean the data
```{r}
homicide_df =
  raw_data %>%
  janitor::clean_names() %>%
  
  # create new variables
  # convert "victim_age" to numeric variable
  mutate(
    city_state = str_c(city, state, sep = ","),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)) %>%
  
  relocate(city_state) %>%
  
  # filter out certain observations
  filter(
    !(city_state %in% c("Dallas,TX", "Phoenix,AZ", " Kansas City,MO", "Tulsa,AL")) &
    victim_race %in% c("Black", "White"))
```

### Model Fitting

#### Baltimore Model Fitting

* Fit a logistic regression model for "Baltimore,MD" and extract estimated OR and the corresponding CIs
```{r}
baltimore_df =
  homicide_df %>%
  filter(city_state == "Baltimore,MD" )

baltimore_fit = 
  glm(resolved ~ victim_age + victim_sex + victim_race, family = binomial, data = baltimore_df) %>%
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    conf.lower = exp(estimate - 1.96*std.error),
    conf.upper = exp(estimate + 1.96*std.error)) %>%
  filter(term == "victim_sexMale") %>%
  select(term, OR, conf.lower, conf.upper)

baltimore_fit
```
The estimate value of the adjusted odds ratio is `r baltimore_fit$OR`, the corresponding confidence interval is [`r baltimore_fit$conf.lower`,`r baltimore_fit$conf.upper`].

#### Iterate over each of the cities

* Build a function that takes city name as input and returns estimated OR and CIs.
```{r}
OR_function = function(city_df) {
  
  city_fit = 
    glm(resolved ~ victim_age + victim_sex + victim_race, family = binomial, data = city_df) %>%
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    conf.lower = exp(estimate - 1.96*std.error),
    conf.upper = exp(estimate + 1.96*std.error)) %>%
  filter(term == "victim_sexMale") %>%
  select(OR, conf.lower, conf.upper)
  
  return(city_fit)
}
```

* Apply the function to each of the cities
```{r}
results_df = 
  homicide_df %>%
  nest(data = uid:resolved) %>%
  mutate(
    fit_results = map(data, OR_function)) %>%
  select(-data) %>%
  unnest(fit_results)

results_df %>%
  knitr::kable(digits = 3)
```

* Generate the plot of estimated ORs and CIs across each city
```{r}
results_df %>%
  mutate(
    city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.lower, ymax = conf.upper)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  
```

### Problem 3

#### Data Cleaning

* Load the data
```{r}
birthwt_raw = read_csv("birthweight.csv")
```

* Clean the data
```{r}
birthwt_df =
  birthwt_raw %>%
  janitor::clean_names() %>%
  
  # convert categorical variables to factors
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) %>%
  
  # drop the missing data
  drop_na()
```

#### Model Fitting

Here I fit a model through forward selection: start with no predictors in the model and add new predictors one by one until no significant predictors can be selected. 
```{r}
min.model = lm(bwt ~ 1, data = birthwt_df)

mod_forward = step(min.model, direction = 'forward', scope = (~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain))
```

The resulting model contains the following predictors: head circumference, birth length, mother's race, mother's weight, gestational age, average number of cigarettes smoked, mother's BMI, baby sex, number of prior live birth, mother's pre-preganacy weight, monthly family income.
  
* Final model
```{r}
mod1 = lm(bwt ~ bhead + blength + mrace + delwt + gaweeks + smoken + ppbmi + 
    babysex + parity + ppwt + fincome, data = birthwt_df)
```

* Plot of model residuals against fitted values
```{r}
birthwt_df %>%
  add_residuals(mod1) %>%
  add_predictions(mod1) %>%
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  labs(title = "Residuals vs. Fitted Values")
```

* Fit two other models for comparison
```{r}
mod2 = lm(bwt ~ blength + gaweeks, data = birthwt_df)
mod3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthwt_df)
```

### Model Comparison
```{r}
cv_df = 
  crossv_mc(birthwt_df, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

* Fit models to training dataset and obtain corresponding RMSEs for the testing data
```{r}
cv_df = 
  cv_df %>%
  mutate(
    forward_mod = map(train, 
                      ~lm(bwt ~ bhead + blength + mrace + delwt + gaweeks + smoken + 
                           ppbmi + babysex + parity + ppwt + fincome, data = .x)),
    main_mod = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    interaction_mod = map(train, ~lm(bwt ~ bhead + blength + babysex + 
                                          bhead*blength + bhead*babysex + blength*babysex + 
                                          bhead*blength*babysex, data = .x))) %>%
  mutate(
    rmse_forward = map2_dbl(forward_mod, test, ~rmse(model = .x, data = .y)),
    rmse_main = map2_dbl(main_mod, test, ~rmse(model = .x, data = .y)),
    rmse_interaction = map2_dbl(interaction_mod, test, ~rmse(model = .x, data = .y)))
  
```

* Plot the prediction error distribution for each candidate model
```{r}
cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

We can conclude from the plot: the forward_selection_model has the lowest RMSE, suggesting it has the highest prediction accuracy thus the best performance. The interaction model has a slightly greater RMSE, but it is more complex compared to simple linear regression model. While the second model (main_effect model) has a very large RMSE.